---
title: "Monthly Canadian Election Poll"
author: "Cesar Y. Villarreal Guzman"
date: "03/10/2020"
output: pdf_document
---

```{r, include=FALSE}
# bibliography: bib.bibtex
## Libraries and initial values
library(tidyverse)

data <- read.csv("survey_data.csv")
```


# Executive Summary
In this experiment we conduct a simulation based study on a survey which aims
to quantify the impact of the second wave of the COVID-19 pandemic on the
political views of Canadian's. 

# Introduction


# Survey Methodology
The objective of the survey is to analyze the impact of the incoming COVID-19
second wave on Canada's population political views. 

The population to be sampled will be the Canadian population. Provided access to
a list of addresses and corresponding emails which will be used as a frame for
this experiment, we generate $\mathbf{S}_h$ by stratified simple random 
sampling (SRSWOR).

```{r, echo=FALSE}
knitr::kable(table(data$Province), caption = "Stratum Sizes", align = "lr",
      col.names = c("Province", "Size"))
```


We seek a 95% confidence accuracy in this study. 


# Results and Discussion

# References

# Apendix

```{r, include =FALSE}
#### CLEAN UP
rm(list = ls())
```

